% Sample University of Calgary Thesis
% This file contains the FRONT MATTER other thna the title page

\chapter{Abstract}

In 1909, in his \textit{Logic Notebook}, Charles Sanders Peirce conducted what, to my knowledge, are the first experiments with many-valued post-Aristotelian logic. As these experiments are entirely contained in a handful of pages in his notebook, which were not published or discussed by other authors during his lifetime, his reasons are open to conjecture. By examining and transcribing these pages and connecting them to his larger body of philosophy, I show how his motivations lie within his views on modality, continuity, and his hypothetical cosmology.


\chapter{Preface}

All of the images contained in this thesis come from the \textit{Charles S. Peirce Papers} collection in the Houghton Library at Harvard. These images have been made freely available as part of the Houghton Library's \textit{Digitization Initiative}. Most of the primary sources for this project come from \textit{The Collected Papers of Charles Sanders Peirce. Electronic Edition}. Much of the information used in this thesis comes from the articles ``Peirce's Triadic Logic", by Max Fisch and Atwell Turquette \citeyearpar{fisch_peirces_1966}, and ``Peirce's Triadic Logic Re-visited", by Robert Lane \citeyearpar{lane_peirces_1999}. The diagram in Chapter 2, Section 2, is a modified version of a diagram presented by Turquette in \citeyear{turquette_dualism_1972}.

This thesis is original, unpublished, independent work by the author,
Brent C. Odland.
  
\chapter{Acknowledgments}


I would like to thank my supervisor, Mark Migotti, for guiding me through the whole process of writing this thesis and patiently listening to the crazy ideas I brought to him before settling on this topic. I would also like to thank Richard Zach, who went far above and beyond the ``duties" of an unofficial co-supervisor throughout the course of this project. Thank you to Nicole Wyatt for reading this thesis and offering helpful suggestions toward its improvement. Thank you also to Jen Smith, for facilitating all of the administrative work that goes into a project like this. Thank you to Anne Levey, for all your support and advice throughout the entirety of my MA. And thanks to my wife, Emma Ritchie, for her constant support and encouragement.

\dedication{For the poor tortoise, who had to wait while Achilles took all his teeny tiny little steps.}

\tableofcontents

\chapter{List of Abbreviations}

\begin{center}
\begin{tabular}{ c c } 

 AFC & \textit{A Fourth Curiosity} \\ 
 CCL & \textit{Cambridge Conference Lectures} \\ 
 CP & \textit{Collected Papers of Charles Sanders Peirce} \\
 EP & \textit{Essential Peirce} \\ 
 L & Refers to letter numbers from \textit{The Robin Catalogue}\\ 
 MS & Refers to manuscript numbers from \textit{The Robin Catalogue}\\ 
 NEM & \textit{New Elements of Mathematics} \cite{peirce1976new} \\ 
 PEM & Principle of excluded middle \\ 
 PC & Principle of non-contradiction \\ 
\end{tabular}
\end{center}

% If you have no tables, delete the next line

%\listoftables

% If you have no figures, delete the next line

%\listoffigures

% Consult Ch 9 of the memoir class manual on how to set up other
% content lists. Note that memoir does not automatically clear the
% page for these. ucalgmthesis fixes this for the default table of
% contents and lists of tables and figures, but not for anything you
% define
